
# Reproducible Research: Assignment 1
## Loading and preprocessing the data

During this step I load the data and delete incomplete rows.
``` {r echo = TRUE}
activity <- read.csv("./activity.csv")
activity_without_na <- activity[complete.cases(activity),]
```

##Calculating the mean total number of steps taken per day
During this step i use dlyr package to calculate sums of steps for each day.

``` {r echo = TRUE}
library(plyr)
steps_per_day <- ddply(activity_without_na, .(date), summarise, steps=sum(steps))
hist(steps_per_day$steps, xlab = "steps", main = "Histogram of the total number of steps taken each day")
mn <-  mean(steps_per_day$steps)
md <- median(steps_per_day$steps)
```
I found that the mean is `r mn` and the median is `r md`

##Finding the average daily activity pattern
During this step i use dlyr package to calculate the mean number of steps for each time in day averaged across all days.

``` {r echo = TRUE}
steps_each_time <- ddply(activity_without_na, .(interval), summarize, steps=mean(steps))
plot(steps_each_time, type ="l")
title("The average number of steps during a day")
interval_number <- steps_each_time[which(steps_each_time$steps == max(steps_each_time$steps)),1]
```
The interval number of maximum number of paces is `r interval_number` 

##Imputing missing values
Now I calculate total number of rows with missing data and fill missing data values with average value of paces for that 5 min interval

```{r echo = TRUE}
library(ggplot2)
number_of_na_rows <- nrow(activity) - nrow(activity_without_na)
filled_activity <- data.frame()

for (i in 1:nrow(activity)){
    if (is.na(activity[i, 1])) {
        temp <- cbind(subset(steps_each_time, interval == activity[i, 3])[2], activity[i, 2], activity[i, 3])
        colnames(temp) <- c("steps", "date", "interval")
        filled_activity <- rbind(filled_activity, temp)
    } 
    else {
        filled_activity <- rbind(filled_activity, activity[i,])
    }
}

steps_per_day_after_filling_na <- ddply(filled_activity, .(date), summarise, steps=sum(steps))
steps_per_day_after_filling_na$type <- "after adding means"
steps_per_day$type <- "before adding means"
two_hists <- rbind(steps_per_day, steps_per_day_after_filling_na)
ggplot(two_hists, aes(steps, fill = type)) + geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity', binwidth = 5000)
mn_after_filling <-  mean(steps_per_day_after_filling_na$steps)
md_after_filling <- median(steps_per_day_after_filling_na$steps)
```
The total number of *NA* is `r number_of_na_rows`
Obtained histogram shows that after adding means the part of movements in the afternoon hours is increased.
The statistical characteristics of the distribution:
mean is equal `r mn_after_filling`
median is equal `r md_after_filling`


## Investigation of the differences in activity patterns between weekdays and weekends

``` {r echo = TRUE}
weekdays_and_weekends <- cbind(filled_activity, weekdays(as.Date(as.character(filled_activity$date))))

for (i in 1:nrow(weekdays_and_weekends)){
    weekdays_and_weekends[i, 4]
    if (as.character(weekdays_and_weekends[i, 4]) == "Saturday" || as.character(weekdays_and_weekends[i, 4]) == "Sunday") {
        weekdays_and_weekends[i, 5] <- "weekend"
        } 
    else {
        weekdays_and_weekends[i, 5] <- "weekday"
    }
}

weekdays_and_weekends[,5] <- as.factor(weekdays_and_weekends[,5])
steps_each_time_filled <- ddply(weekdays_and_weekends, .(interval, V5), summarize, steps=mean(steps))
ggplot(data=steps_each_time_filled, aes(x=interval, y=steps, colour=V5)) + geom_line() + ggtitle("The average number of steps during a day")
```

This plot indicates that the person's activity is inscreased in the morning (he/she goes to work) then in is decreased in comparison with weekend which indicates that this person has a job with low physical activity.

